Full-day, in home validation of infant body position measurements from inertial sensors

### 2023-03-14 manuscript.Rmd
R citations

Infants were recruited in one of two age groups: *Younger* infants were between 4 and 7 months and *Older* infants were between 11 and 14 months. There were `r num_younger` infants in the 4-7 month group (X female) and `r num_older` in the 11-14 month group (X female). Families were recruited through social media advertisements and from community events in Southern California. The ethnicity for infants was reported as X Hispanix/Latinx and X for not Hispanix/Latinx. Race was reported as X Caucasian, X Asian, X ~~~ and X others. Parents were compensated \$30 in cash for each visit they completed. The BLINDED Institutional Review Board reviewed and approved all procedures associated with the study. All participants gave their informed consent to participate.
<!-- TODO RACE ETHNICITY SEX, COUNT BY HAND -->

Most participants were tested in a single session (*n* = `r nrow(num_sessions %>% filter(n == 1))`), but `r nrow(num_sessions %>% filter(n > 1))` participants contributed between 2-4 sessions as part of an ongoing longitudinal study. Only 1 session was excluded due to a technical error---one of the four IMU sensors failed to record, resulting in an unusable set of data for classification. Across the two age groups, we report data on a total of `r length(unique(ppts$id_uni))` sessions, with `r num_sessions %>% filter(group == "Younger") %>% pull(n) %>% sum` sessions from younger infants and `r num_sessions %>% filter(group == "Older") %>% pull(n) %>% sum` sessions from older infants. Across sessions, younger infants' age ranged from `r age["Younger","min"]` to `r age["Younger","max"]` (*M* = `r age["Younger","mean"]`) and older infants' age ranged from `r age["Older","min"]` to `r age["Older","max"]` (*M* = `r age["Older","mean"]`).

<!-- with X including permission for data to be shared on Databrary.  -->

@@

Four inertial movement units (IMUs) were used to record infant movement across the day (MC10 Biostamp). A custom garment was made to hold the IMUs: Internal pockets were sewn into a snug-fitting pair of infant leggings so that IMUs would stay close to the body and so that infants could not pull out the sensors. On each side of the body there was a pocket over the hip and a pocket just above the ankle. Each sensor recorded accelerometer and gyroscope data at 62.5 Hz throughout the day, with sufficient battery and on-device storage to record for approximately 12 hours. Infants also wore a LENA&reg; recorder throughout the day in the front pocket of a LENA&reg; shirt, located near the infant’s chest.

Video recordings were captured using an action camera on a miniature tripod (Insta360 ONE R) that caregivers placed in the room that the infant was in. The proprietary "Boosted Battery Base" attached to the action camera to allow for a total of 3 hours of recording. However, as described below, this created two video files temporally separated by a gap of approximately 40-45 s. Caregivers also received a log sheet to record times that infants napped as well as times that the sensor garment was removed from the infant (e.g., baths, diaper changes).

@@

```{r exemplar-timeline, out.width="0.99\\linewidth", include=TRUE, fig.align="center", fig.cap="Example Timeline Caption.", echo=FALSE}
knitr::include_graphics("figures/timeline.png")
# Generated from analysis-pt2/methods-exemplar-agreement.R
# Annotated in illustrator based on event times:
# Exemplar participant is 120-1
# > pt2_start_time = "2022-11-13 11:11:00 PST"
# > pt2_end_time = "2022-11-13 12:41:01 PST"
# > session_param$start_time_coded = "2022-11-13 10:02:40 PST"
```

Figure \@ref(fig:exemplar-timeline)) shows an exemplar timeline of the entire procedure and recording periods for a single participant. On the day of the visit, a researcher arrived at the participant’s home in the morning between and prepared all the equipment at the doorstep. In order to synchronize all three recording devices, the researcher began by first turning on the video camera and the LENA audio device (the IMUs were already configured and placed at arrival because they required a proprietary sensor dock to begin recording in the laboratory). To create an easily recognizable synchronization point between the video recording and IMU data, the researcher dropped or struck the leggings (containing the IMUs) on a surface in view of the camera, as in @FranchakScott2021. All the equipment---once recording and with synchronization information recorded---was placed inside a large bucket and left outside the family's front door.

The researcher then called the caregiver on the phone and walked them through a set of procedures needed to properly set up the equipment and record video for classifier training and testing. At the start of this "guided call", the caregiver was instructed to place the camera in an area that captured the majority of the room. Next, they were asked to put the pair of leggings and shirt on their infant, with the researcher providing guidance about how to correctly orient the garments.

@@

Afterwards, they were to go about their day as usual with the infant wearing the sensor garment until their bedtime, only taking off the sensor for naps, baths, and diaper changes. The caregiver logged the times the sensors were removed (blank areas in the timeline in Figure \@ref(fig:exemplar-timeline))) or the child took a nap (gray areas in the timeline in Figure \@ref(fig:exemplar-timeline))) so that those times could be excluded from analysis. The following day a researcher picked up the equipment, verified the paperwork was signed, and compensated the participant.

Because the camera only had the battery life to record for ~3 hours (divided into two 90-minute video files), this divided the day into different periods for analysis. As seen in the bottom of Figure \@ref(fig:exemplar-timeline)), the *video period* comprised the first three hours of recording starting from the researcher's arrival when they turned on the camera. The first 90-minute video file, termed the *proximal comparison*, contained the activities during the guided call followed by a period of infants and caregivers resuming their normal activities. Because this video contained the synchronization point, the data in this period had temporal synchrony between IMU and video data that contained errors of no more than 30-60 ms (1-2 video frames). When the first video file ended, a second video file was recorded, termed the *distal comparison*. This video recorded the next 90 minutes of natural activity. However, because there was a variable gap of ~40 s between the two videos, temporal synchrony in the distal comparison video was approximate containing offsets of ~ 5 s in either direction.
<!-- HOW MANY PPTS DON"T HAVE DISTAL COMPARISON VIDEOS? -->

@@

The proximal and distal comparison videos for each participant were annotated by trained human coders to identify infant *body position* into one of 5 mutually-exclusive categories: supine, prone, sitting, upright, or held by caregiver. All coding was done using Datavyu software (datavyu.org).

Supine was coded when the infant was lying on their back or was reclined up to a 45 degrees angle. Supine was also coded in the rare cases when the infant was laying on their side. Prone was coded when the infant was lying on their stomach, was on all fours in a downward dog position, or was crawling. We scored sitting to include any form of the following seated positions: 1) infants sat with their buttocks on a surface, such as on the floor or a caregiver’s lap, 2) infant was in a kneeling-sit position, in which their knees were on the ground with their legs tucked underneath the buttocks, and 3) infant was in a seating device, such as a high chair, that kept the torso oriented perpendicular to the ground (a reclined position, such as in a young infant's car seat, would be counted as supine). Upright was coded when the infant was standing on the ground with two feet or walking (regardless of whether infants' balance was assisted by a caregiver or with their hands holding onto something for support). When an infant was carried by a caregiver, held was coded. However, when the caregiver was sitting with the infant in their lap the infant’s body position was coded as if the caregiver was a surface (e.g., if the infant was sitting on the caregiver’s lap this was coded as sitting). Times during the video when the infant was out of view were excluded. Periods when the leggings were being adjusted or taken off the infant were also excluded, as were transitions between body positions.

@@

<!-- R packages? Shared code -->
The same machine learning classification process was used as in prior work [@FranchakScott2021]. Using the synchronization point, human-coded body annotations from video were linked to the corresponding times in the IMU timeseries data. A single, merged dataset was created with synchronized accelerometer signals (in three orientations: X, Y, and Z) and gyroscope signals (in three orientations: roll, pitch, and yaw) for each of the four sensors (left hip, right hip, left ankle, right ankle) with the corresponding timestamp and body position code.

@@

Windowed datasets were used for machine learning classification and validation. For each analysis reported in the results, a subset of data were defined as a "training" set and another, independent portion of the data were defined as a "testing" set. Different training/testing set combinations were used to address different goals; here we describe the general procedure of how a classifier was derived from a training set. Random forest models [@Breiman2001] were trained to predict the human coded body position label for each window from the set of 436 motion features using the *randomForest* package in R [@LiawWiener2002]. Random forest models use random subsets of features and random subsets of data across many iterations; this random subsetting creates models that are less likely to be overfit. The *tidymodels* set of packages in R [@tidymodels] was used to tune the "mtry" and "ntrees" hyperparameters, which control the number of columns sampled and the total number of trees created when modeling. We use an mtry parameter of 44 and and ntree parameter of 550, however, changing hyperparameters had little impact on model performance. Just like a linear regression model, the resulting random forest model could later be applied to a set of testing data with the *predict* function. Validation analyses below will compare the human-coded annotations and random forest model predictions for testing data that are independent from training data, providing a test of whether models can successfully generalize.

>>>>

Infants were recruited in one of two age groups: *Younger* infants were between 4 and 7 months and *older* infants were between 11 and 14 months. There were `r num_younger` infants in the 4-7 month group (X female) and `r num_older` in the 11-14 month group (X female). Families were recruited through social media advertisements and from community events in Southern California. Caregivers reported the ethnicity of infants as X Hispanix/Latinx and X for not Hispanix/Latinx. Race was reported as X Caucasian, X Asian, X ~~~ and X others. Families were compensated \$30 for every home recording session they completed. The BLINDED Institutional Review Board reviewed and approved all procedures associated with the study. All caregivers gave their informed consent before the start of the study.
<!-- TODO RACE ETHNICITY SEX, COUNT BY HAND -->

Most participants were tested in a single session (*n* = `r nrow(num_sessions %>% filter(n == 1))`), but `r nrow(num_sessions %>% filter(n > 1))` participants contributed between 2-4 sessions as part of an ongoing longitudinal study. Only 1 session was excluded due to a technical error---one of the four IMU sensors failed to record, resulting in an unusable set of data for classification. Across the two age groups, we report data on a total of `r length(unique(ppts$id_uni))` sessions, with `r num_sessions %>% filter(group == "Younger") %>% pull(n) %>% sum` sessions from younger infants and `r num_sessions %>% filter(group == "Older") %>% pull(n) %>% sum` sessions from older infants. Across sessions, younger infants' age ranged from `r apa_num(age["Younger","min"], digits = 1)` to `r apa_num(age["Younger","max"], digits = 1)` months (*M* = `r apa_num(age["Younger","mean"], digits = 1)`) and older infants' age ranged from `r apa_num(age["Older","min"], digits = 1)` to `r apa_num(age["Older","max"], digits = 1)` months (*M* = `r apa_num(age["Older","mean"], digits = 1)`) .

@@

Four inertial movement units (IMUs) were used to record infant movement across the day (MC10 Biostamp). A custom garment was designed to hold the IMUs: Internal pockets were sewn into a snug-fitting pair of infant leggings so that IMUs would stay close to the body (reducing vibration), which also prevented infants from pulling out the sensors. A pocket over the hip and a pocket just above the ankle were sewn on the outside of the right and left legs of the garment. Each sensor recorded accelerometer and gyroscope data at 62.5 Hz throughout the day, with sufficient battery and onboard storage to record for approximately 12 hours. Infants also wore a LENA&reg; recorder throughout the day in the front pocket of a LENA&reg; shirt, located near the infant’s chest.

Video recordings were captured using an action camera on a miniature tripod (Insta360 ONE R) that caregivers placed in the room that the infant was in. The proprietary "Boosted Battery Base" attached to the action camera to allow for a total of 3 hours of recording. As described below, the video recording was not continuous. The camera created two video files temporally separated by a gap of approximately 40-45 s. Caregivers also received a log sheet to record times that infants napped as well as times that the sensor garment was removed from the infant (e.g., baths, diaper changes).

@@

```{r exemplar-timeline, out.width="0.99\\linewidth", include=TRUE, fig.align="center", fig.cap="Timeline from an exemplar participant (older infant 15). The top row shows the model-predicted body position across the entire recording period. Annotations indicate when the video camera was turned on by the experimenter when arriving at the house, when the sensors were first placed on the infant, when the guided call took place, and when the video files were recorded. Gray areas on the timeline indicate naps, and white areas indicate times when the sensors were removed. The bottom row shows a zoomed-in view of the video period during which ground truth data were available. The top timeline shows human-coded body position and the bottom row shows model-predicted body position; these were the data used for validation. The first part of the video period was the proximal comparison, when video and motion data were highly synchronized. The second part of the video recording was the distal comparison that had coarser synchronization. Accuracy data for the distal comparison are provided overall and during 10-minute bins, marked by vertical dashed lines.", echo=FALSE}
knitr::include_graphics("figures/timeline.png")
# Generated from analysis-pt2/methods-exemplar-agreement.R
# Annotated in illustrator based on event times:
# Exemplar participant is 120-1
# > pt2_start_time = "2022-11-13 11:11:00 PST"
# > pt2_end_time = "2022-11-13 12:41:01 PST"
# > session_param$start_time_coded = "2022-11-13 10:02:40 PST"
```

Figure \@ref(fig:exemplar-timeline)) shows an exemplar timeline of the entire procedure and recording periods for a single participant. On the day of the visit, a researcher arrived at the participant’s home in the morning between and prepared all the equipment at the doorstep. In order to synchronize all three recording devices, the researcher began by first turning on the video camera and the LENA audio device (the IMUs were already configured and placed before arrival because they required a proprietary sensor dock to begin recording in the laboratory). To create an easily recognizable synchronization point between the video recording and IMU data, the researcher dropped or struck the leggings (containing the IMUs) on a surface in view of the camera, as in @FranchakScott2021. All the equipment---once recording and with synchronization information recorded---was placed inside a large bucket and left outside the family's front door. The researcher then called the caregiver on the phone and walked them through a set of procedures needed to properly set up the equipment and record video for ground truth human annotation of body position. At the start of this "guided call", the caregiver was instructed to place the camera in an area that captured the majority of the room. Next, they were asked to put the pair of leggings and shirt on their infant, with the researcher providing guidance about how to correctly orient the garments.

@@

Afterwards, they were asked to go about their day as usual with the infant wearing the sensor garment until their bedtime, only taking off the sensor for naps, baths, and diaper changes. The caregiver logged the times the sensors were removed (blank areas in the timeline in Figure \@ref(fig:exemplar-timeline))) or the child took a nap (gray areas in the timeline in Figure \@ref(fig:exemplar-timeline))) so that those times could be excluded from analysis. The following day a researcher picked up the equipment, verified that the paperwork was signed, and compensated the participant.

Because the camera only had the battery life to record for ~3 hours (split into two 90-minute video files), this divided the day into different periods for analysis. As seen in the bottom of Figure \@ref(fig:exemplar-timeline)), the *video period* comprised the first three hours of recording starting from the researcher's arrival when they turned on the camera. The first 90-minute video file, termed the *proximal comparison*, contained the activities during the guided call followed by a period of infants and caregivers resuming their normal activities. Because this video contained the synchronization point, the data in this period had high temporal synchrony between IMU and video data. Synchrony errors were less than 30-60 ms (1-2 video frames). When the first video file ended, a second video file was recorded, termed the *distal comparison*. This video recorded the next 90 minutes of natural activity. However, because there was a variable gap of ~40 s between the two videos, temporal synchrony in the distal comparison video was approximate containing offsets of ~ 5 s in either direction.

@@

The proximal and distal comparison videos for each participant were annotated by trained human coders to identify infant *body position* into one of 5 mutually-exclusive categories following the definitions in prior work [@FranchakScott2021]: supine, prone, sitting, upright, or held by caregiver. All coding was done using Datavyu software (datavyu.org).

Supine was coded when the infant was lying on their back or was reclined up to a 45 degree angle. Supine was also coded in the rare cases when the infant was laying on their side. Prone was coded when the infant was lying on their stomach, was stationary supported by the hands/knees or the hands/feet, or was crawling. We scored sitting to include any form of the following seated positions: 1) infants sat with their buttocks on a surface, such as on the floor or a caregiver’s lap, 2) infant was in a kneeling-sit position, in which their knees were on the ground with their legs tucked underneath the buttocks, and 3) infant was in a seating device, such as a high chair, that kept the torso oriented perpendicular to the ground (a reclined position, such as in a young infant's car seat, would be counted as supine). Upright was coded when the infant was standing on the ground with two feet or walking (regardless of whether infants' balance was assisted by a caregiver or with their hands holding onto something for support). When an infant was carried by a caregiver, held was coded. However, when the caregiver was sitting with the infant in their lap the infant’s body position was coded as if the caregiver was a surface (e.g., if the infant was sitting on the caregiver’s lap this was coded as sitting). Times during the video when the infant was out of view were excluded. Periods when the leggings were being adjusted or taken off the infant were also excluded, as were transitions between body positions.

@@

The same machine learning classification process was used as in prior work [@FranchakScott2021]. Using the synchronization point, human-coded body annotations from video were linked to the corresponding times in the IMU timeseries data. A single, merged dataset was created with synchronized accelerometer signals (in three orientations: X, Y, and Z) and gyroscope signals (in three orientations: roll, pitch, and yaw) for each of the four sensors (left hip, right hip, left ankle, right ankle) with the corresponding timestamp and body position code.

@@

Windowed datasets were used for machine learning classification and validation. For each analysis reported in the results, a subset of data were defined as a "training" set and another, independent portion of the data were defined as a "testing" set. Different training/testing set combinations were used to address different goals; here we describe the general procedure of how a classifier was derived from a training set. Random forest models [@Breiman2001] were trained to predict the human coded body position label for each window from the set of 436 motion features using the *randomForest* package [@LiawWiener2002] in R version 4.1.2 [@R42]. Random forest models use random subsets of features and random subsets of data across many iterations; this random subsetting creates models that are less likely to be overfit. The *tidymodels* set of packages in R [@tidymodels] was used to tune the "mtry" and "ntrees" hyperparameters, which control the number of columns sampled and the total number of trees created when modeling. We use an mtry parameter of 44 and and ntree parameter of 550, however, changing hyperparameters had little impact on model performance. Just like a linear regression model, the resulting random forest model could later be applied to a set of testing data with the *predict* function. Validation analyses below will compare the human-coded annotations and random forest model predictions for testing data that are independent from training data, providing a test of whether models can successfully generalize.

### 2023-03-14 manuscript.tex
R citations

Infants were recruited in one of two age groups: \emph{Younger} infants were between 4 and 7 months and \emph{Older} infants were between 11 and 14 months. There were 8 infants in the 4-7 month group (X female) and 14 in the 11-14 month group (X female). Families were recruited through social media advertisements and from community events in Southern California. The ethnicity for infants was reported as X Hispanix/Latinx and X for not Hispanix/Latinx. Race was reported as X Caucasian, X Asian, X \textasciitilde\textasciitilde\textasciitilde{} and X others. Parents were compensated \$30 in cash for each visit they completed. The BLINDED Institutional Review Board reviewed and approved all procedures associated with the study. All participants gave their informed consent to participate.

Most participants were tested in a single session (\emph{n} = 16), but 6 participants contributed between 2-4 sessions as part of an ongoing longitudinal study. Only 1 session was excluded due to a technical error---one of the four IMU sensors failed to record, resulting in an unusable set of data for classification. Across the two age groups, we report data on a total of 33 sessions, with 13 sessions from younger infants and 20 sessions from older infants. Across sessions, younger infants' age ranged from 3.84 to 7.16 (\emph{M} = 5.03) and older infants' age ranged from 10.74 to 14.23 (\emph{M} = 11.75).

@@

Four inertial movement units (IMUs) were used to record infant movement across the day (MC10 Biostamp). A custom garment was made to hold the IMUs: Internal pockets were sewn into a snug-fitting pair of infant leggings so that IMUs would stay close to the body and so that infants could not pull out the sensors. On each side of the body there was a pocket over the hip and a pocket just above the ankle. Each sensor recorded accelerometer and gyroscope data at 62.5 Hz throughout the day, with sufficient battery and on-device storage to record for approximately 12 hours. Infants also wore a LENA® recorder throughout the day in the front pocket of a LENA® shirt, located near the infant's chest.

Video recordings were captured using an action camera on a miniature tripod (Insta360 ONE R) that caregivers placed in the room that the infant was in. The proprietary ``Boosted Battery Base'' attached to the action camera to allow for a total of 3 hours of recording. However, as described below, this created two video files temporally separated by a gap of approximately 40-45 s. Caregivers also received a log sheet to record times that infants napped as well as times that the sensor garment was removed from the infant (e.g., baths, diaper changes).

@@

\caption{Example Timeline Caption.}\label{fig:exemplar-timeline}
\end{figure}

Figure \ref{fig:exemplar-timeline}) shows an exemplar timeline of the entire procedure and recording periods for a single participant. On the day of the visit, a researcher arrived at the participant's home in the morning between and prepared all the equipment at the doorstep. In order to synchronize all three recording devices, the researcher began by first turning on the video camera and the LENA audio device (the IMUs were already configured and placed at arrival because they required a proprietary sensor dock to begin recording in the laboratory). To create an easily recognizable synchronization point between the video recording and IMU data, the researcher dropped or struck the leggings (containing the IMUs) on a surface in view of the camera, as in Franchak et al. (2021). All the equipment---once recording and with synchronization information recorded---was placed inside a large bucket and left outside the family's front door.

The researcher then called the caregiver on the phone and walked them through a set of procedures needed to properly set up the equipment and record video for classifier training and testing. At the start of this ``guided call'', the caregiver was instructed to place the camera in an area that captured the majority of the room. Next, they were asked to put the pair of leggings and shirt on their infant, with the researcher providing guidance about how to correctly orient the garments.

@@

Afterwards, they were to go about their day as usual with the infant wearing the sensor garment until their bedtime, only taking off the sensor for naps, baths, and diaper changes. The caregiver logged the times the sensors were removed (blank areas in the timeline in Figure \ref{fig:exemplar-timeline})) or the child took a nap (gray areas in the timeline in Figure \ref{fig:exemplar-timeline})) so that those times could be excluded from analysis. The following day a researcher picked up the equipment, verified the paperwork was signed, and compensated the participant.

Because the camera only had the battery life to record for \textasciitilde3 hours (divided into two 90-minute video files), this divided the day into different periods for analysis. As seen in the bottom of Figure \ref{fig:exemplar-timeline}), the \emph{video period} comprised the first three hours of recording starting from the researcher's arrival when they turned on the camera. The first 90-minute video file, termed the \emph{proximal comparison}, contained the activities during the guided call followed by a period of infants and caregivers resuming their normal activities. Because this video contained the synchronization point, the data in this period had temporal synchrony between IMU and video data that contained errors of no more than 30-60 ms (1-2 video frames). When the first video file ended, a second video file was recorded, termed the \emph{distal comparison}. This video recorded the next 90 minutes of natural activity. However, because there was a variable gap of \textasciitilde40 s between the two videos, temporal synchrony in the distal comparison video was approximate containing offsets of \textasciitilde{} 5 s in either direction.

@@

The proximal and distal comparison videos for each participant were annotated by trained human coders to identify infant \emph{body position} into one of 5 mutually-exclusive categories: supine, prone, sitting, upright, or held by caregiver. All coding was done using Datavyu software (datavyu.org).

Supine was coded when the infant was lying on their back or was reclined up to a 45 degrees angle. Supine was also coded in the rare cases when the infant was laying on their side. Prone was coded when the infant was lying on their stomach, was on all fours in a downward dog position, or was crawling. We scored sitting to include any form of the following seated positions: 1) infants sat with their buttocks on a surface, such as on the floor or a caregiver's lap, 2) infant was in a kneeling-sit position, in which their knees were on the ground with their legs tucked underneath the buttocks, and 3) infant was in a seating device, such as a high chair, that kept the torso oriented perpendicular to the ground (a reclined position, such as in a young infant's car seat, would be counted as supine). Upright was coded when the infant was standing on the ground with two feet or walking (regardless of whether infants' balance was assisted by a caregiver or with their hands holding onto something for support). When an infant was carried by a caregiver, held was coded. However, when the caregiver was sitting with the infant in their lap the infant's body position was coded as if the caregiver was a surface (e.g., if the infant was sitting on the caregiver's lap this was coded as sitting). Times during the video when the infant was out of view were excluded. Periods when the leggings were being adjusted or taken off the infant were also excluded, as were transitions between body positions.

@@

Windowed datasets were used for machine learning classification and validation. For each analysis reported in the results, a subset of data were defined as a ``training'' set and another, independent portion of the data were defined as a ``testing'' set. Different training/testing set combinations were used to address different goals; here we describe the general procedure of how a classifier was derived from a training set. Random forest models (Breiman, 2001) were trained to predict the human coded body position label for each window from the set of 436 motion features using the \emph{randomForest} package in R (Liaw et al., 2002). Random forest models use random subsets of features and random subsets of data across many iterations; this random subsetting creates models that are less likely to be overfit. The \emph{tidymodels} set of packages in R (Kuhn \& Wickham, 2020) was used to tune the ``mtry'' and ``ntrees'' hyperparameters, which control the number of columns sampled and the total number of trees created when modeling. We use an mtry parameter of 44 and and ntree parameter of 550, however, changing hyperparameters had little impact on model performance. Just like a linear regression model, the resulting random forest model could later be applied to a set of testing data with the \emph{predict} function. Validation analyses below will compare the human-coded annotations and random forest model predictions for testing data that are independent from training data, providing a test of whether models can successfully generalize.

@@

>>>>

Infants were recruited in one of two age groups: \emph{Younger} infants were between 4 and 7 months and \emph{older} infants were between 11 and 14 months. There were 8 infants in the 4-7 month group (X female) and 14 in the 11-14 month group (X female). Families were recruited through social media advertisements and from community events in Southern California. Caregivers reported the ethnicity of infants as X Hispanix/Latinx and X for not Hispanix/Latinx. Race was reported as X Caucasian, X Asian, X \textasciitilde\textasciitilde\textasciitilde{} and X others. Families were compensated \$30 for every home recording session they completed. The BLINDED Institutional Review Board reviewed and approved all procedures associated with the study. All caregivers gave their informed consent before the start of the study.

Most participants were tested in a single session (\emph{n} = 16), but 6 participants contributed between 2-4 sessions as part of an ongoing longitudinal study. Only 1 session was excluded due to a technical error---one of the four IMU sensors failed to record, resulting in an unusable set of data for classification. Across the two age groups, we report data on a total of 33 sessions, with 13 sessions from younger infants and 20 sessions from older infants. Across sessions, younger infants' age ranged from 3.8 to 7.2 months (\emph{M} = 5.0) and older infants' age ranged from 10.7 to 14.2 months (\emph{M} = 11.7) .

@@

Four inertial movement units (IMUs) were used to record infant movement across the day (MC10 Biostamp). A custom garment was designed to hold the IMUs: Internal pockets were sewn into a snug-fitting pair of infant leggings so that IMUs would stay close to the body (reducing vibration), which also prevented infants from pulling out the sensors. A pocket over the hip and a pocket just above the ankle were sewn on the outside of the right and left legs of the garment. Each sensor recorded accelerometer and gyroscope data at 62.5 Hz throughout the day, with sufficient battery and onboard storage to record for approximately 12 hours. Infants also wore a LENA® recorder throughout the day in the front pocket of a LENA® shirt, located near the infant's chest.

Video recordings were captured using an action camera on a miniature tripod (Insta360 ONE R) that caregivers placed in the room that the infant was in. The proprietary ``Boosted Battery Base'' attached to the action camera to allow for a total of 3 hours of recording. As described below, the video recording was not continuous. The camera created two video files temporally separated by a gap of approximately 40-45 s. Caregivers also received a log sheet to record times that infants napped as well as times that the sensor garment was removed from the infant (e.g., baths, diaper changes).

@@

\caption{Timeline from an exemplar participant (older infant 15). The top row shows the model-predicted body position across the entire recording period. Annotations indicate when the video camera was turned on by the experimenter when arriving at the house, when the sensors were first placed on the infant, when the guided call took place, and when the video files were recorded. Gray areas on the timeline indicate naps, and white areas indicate times when the sensors were removed. The bottom row shows a zoomed-in view of the video period during which ground truth data were available. The top timeline shows human-coded body position and the bottom row shows model-predicted body position; these were the data used for validation. The first part of the video period was the proximal comparison, when video and motion data were highly synchronized. The second part of the video recording was the distal comparison that had coarser synchronization. Accuracy data for the distal comparison are provided overall and during 10-minute bins, marked by vertical dashed lines.}\label{fig:exemplar-timeline}
\end{figure}

Figure \ref{fig:exemplar-timeline}) shows an exemplar timeline of the entire procedure and recording periods for a single participant. On the day of the visit, a researcher arrived at the participant's home in the morning between and prepared all the equipment at the doorstep. In order to synchronize all three recording devices, the researcher began by first turning on the video camera and the LENA audio device (the IMUs were already configured and placed before arrival because they required a proprietary sensor dock to begin recording in the laboratory). To create an easily recognizable synchronization point between the video recording and IMU data, the researcher dropped or struck the leggings (containing the IMUs) on a surface in view of the camera, as in Franchak et al. (2021). All the equipment---once recording and with synchronization information recorded---was placed inside a large bucket and left outside the family's front door. The researcher then called the caregiver on the phone and walked them through a set of procedures needed to properly set up the equipment and record video for ground truth human annotation of body position. At the start of this ``guided call'', the caregiver was instructed to place the camera in an area that captured the majority of the room. Next, they were asked to put the pair of leggings and shirt on their infant, with the researcher providing guidance about how to correctly orient the garments.

@@

Afterwards, they were asked to go about their day as usual with the infant wearing the sensor garment until their bedtime, only taking off the sensor for naps, baths, and diaper changes. The caregiver logged the times the sensors were removed (blank areas in the timeline in Figure \ref{fig:exemplar-timeline})) or the child took a nap (gray areas in the timeline in Figure \ref{fig:exemplar-timeline})) so that those times could be excluded from analysis. The following day a researcher picked up the equipment, verified that the paperwork was signed, and compensated the participant.

Because the camera only had the battery life to record for \textasciitilde3 hours (split into two 90-minute video files), this divided the day into different periods for analysis. As seen in the bottom of Figure \ref{fig:exemplar-timeline}), the \emph{video period} comprised the first three hours of recording starting from the researcher's arrival when they turned on the camera. The first 90-minute video file, termed the \emph{proximal comparison}, contained the activities during the guided call followed by a period of infants and caregivers resuming their normal activities. Because this video contained the synchronization point, the data in this period had high temporal synchrony between IMU and video data. Synchrony errors were less than 30-60 ms (1-2 video frames). When the first video file ended, a second video file was recorded, termed the \emph{distal comparison}. This video recorded the next 90 minutes of natural activity. However, because there was a variable gap of \textasciitilde40 s between the two videos, temporal synchrony in the distal comparison video was approximate containing offsets of \textasciitilde{} 5 s in either direction.

@@

The proximal and distal comparison videos for each participant were annotated by trained human coders to identify infant \emph{body position} into one of 5 mutually-exclusive categories following the definitions in prior work (Franchak et al., 2021): supine, prone, sitting, upright, or held by caregiver. All coding was done using Datavyu software (datavyu.org).

Supine was coded when the infant was lying on their back or was reclined up to a 45 degree angle. Supine was also coded in the rare cases when the infant was laying on their side. Prone was coded when the infant was lying on their stomach, was stationary supported by the hands/knees or the hands/feet, or was crawling. We scored sitting to include any form of the following seated positions: 1) infants sat with their buttocks on a surface, such as on the floor or a caregiver's lap, 2) infant was in a kneeling-sit position, in which their knees were on the ground with their legs tucked underneath the buttocks, and 3) infant was in a seating device, such as a high chair, that kept the torso oriented perpendicular to the ground (a reclined position, such as in a young infant's car seat, would be counted as supine). Upright was coded when the infant was standing on the ground with two feet or walking (regardless of whether infants' balance was assisted by a caregiver or with their hands holding onto something for support). When an infant was carried by a caregiver, held was coded. However, when the caregiver was sitting with the infant in their lap the infant's body position was coded as if the caregiver was a surface (e.g., if the infant was sitting on the caregiver's lap this was coded as sitting). Times during the video when the infant was out of view were excluded. Periods when the leggings were being adjusted or taken off the infant were also excluded, as were transitions between body positions.

@@

Windowed datasets were used for machine learning classification and validation. For each analysis reported in the results, a subset of data were defined as a ``training'' set and another, independent portion of the data were defined as a ``testing'' set. Different training/testing set combinations were used to address different goals; here we describe the general procedure of how a classifier was derived from a training set. Random forest models (Breiman, 2001) were trained to predict the human coded body position label for each window from the set of 436 motion features using the \emph{randomForest} package (Liaw et al., 2002) in R version 4.1.2 (R Core Team, 2021). Random forest models use random subsets of features and random subsets of data across many iterations; this random subsetting creates models that are less likely to be overfit. The \emph{tidymodels} set of packages in R (Kuhn \& Wickham, 2020) was used to tune the ``mtry'' and ``ntrees'' hyperparameters, which control the number of columns sampled and the total number of trees created when modeling. We use an mtry parameter of 44 and and ntree parameter of 550, however, changing hyperparameters had little impact on model performance. Just like a linear regression model, the resulting random forest model could later be applied to a set of testing data with the \emph{predict} function. Validation analyses below will compare the human-coded annotations and random forest model predictions for testing data that are independent from training data, providing a test of whether models can successfully generalize.

@@

\leavevmode\vadjust pre{\hypertarget{ref-R42}{}}%
R Core Team. (2021). \emph{R: A language and environment for statistical computing}. {R Foundation for Statistical Computing}.
